### 2020-01-27 report.tex
initial

\section{Related Work}
\subsection{Trail Saving for Assumption-Based SAT}
Trail saving was briefly introduced in the context of assumption-based SAT \cite{DBLP:conf/sat/HickeyB19} in order to speed up the propagation of the assumptions each time the solver backtracks past the assumption levels. When the assumption-based SAT solver backtracks past the assumption levels (i.e., during a restart or whenever a unit clause is learned), the solver will have to repeat much of the work it had previously done when it propagates through the assumptions again. In fact, the set of literals whose truth value is determined by the level 0 implicants and the assumption literals can only grow throughout the remainder of the solving episode, unless a conflict occurs.

In order to take advantage of this fact, we use "trail saving" to store all of the literals assigned during the assumption levels in a cache along with their reason clauses. Before propagating the assumption levels again, this cache is accessed in the same order as it had previously appeared on the trail to assign all of the previous propagants automatically. One must be careful to check that the truth value of the literal is not already assigned before assigning it; if the truth value of the literal is already true then there is no point re-assigning it, and if the truth value of the literal is already false then there is a conflict.

\subsection{Comparison to Chronological Backtracking}
As mentioned earlier, chronological backtracking also achieves the theoretical speedups that trail saving achieves. One advantage of chronological backtracking is that it does not require the caching of the trail or accessing this cache to "restore". It simply leaves a vast portion of the trail in tact rather than backtracking all the way to the conflict level. Chronological backtracking will also speed up propagations slightly more, because it leaves more literals assigned at once which leads to clauses being detected as unit more quickly. One disadvantage that chronological backtracking has compared to trail saving is that it doesn't allow the search as much flexibility. Because chronological backtracking leaves a large portion of the trail assigned, it will behave similarly to how the solver would behave if it was to force the decisions that do occur within this large portion of literals, whereas we do not force any decisions during trail saving. Another disadvantage is that, because propagation is not performed at every decision level, literals that are already assigned do not have a chance to move their decision levels up and this could potentially effect the lbd score or length of the learnt clause that gets generated at the next conflict. An example of this is when the solver was supposed to backtrack over a literal that would end up being unit propagated at level 0, but stays assigned at its current decision level when chronological backtracking is used. Additionally, chronological backtracking breaks the invarant that the all literals appear on the trail in monotonically increasing order by decision level, which makes its implementation slightly more complicated.

>>>>

\section{Related Work}
\subsection{Trail Saving for Assumption-Based SAT}
Trail saving was briefly introduced in the context of assumption-based SAT \cite{DBLP:conf/sat/HickeyB19} in order to speed up the propagation of the assumptions each time the solver backtracked past the assumption levels. Note that the process is the same, but it only works to restore the propagants and implicants of the assumptions and only after a learnt unit clause. Becuase the number of propagants and implicants of the assumptions tends to become dominated by the number of assumptions in the applications it was tested on, the technique did not yield a significant improvement. It would be interesting to combine the extended version of trail saving outlined in this paper with trail saving over the assumptions to see if that could produce better results.

\subsection{Comparison to Chronological Backtracking}
As mentioned earlier, chronological backtracking also achieves some of the theoretical speedups that trail saving achieves. One advantage of chronological backtracking is that it does not require any caching of the trail or accessing such a cache during propagation; it simply leaves a vast portion of the trail intact rather than backtracking all the way to the conflict level. Chronological backtracking will also speed up propagations slightly more than trail saving, because it leaves more literals assigned at once which leads to clauses being detected as unit more quickly. One disadvantage that chronological backtracking has compared to trail saving is that it doesn't allow the search as much flexibility. Because chronological backtracking leaves a large portion of the trail assigned, it will behave similarly to how the solver would behave if it was to force the decisions that do occur within this large portion of literals, whereas trail saving does not force any decisions. Another disadvantage is that, because propagation is not performed at every decision level, literals that are already assigned do not have a chance to move their decision levels up and this could potentially effect the lbd score or length of the learnt clause that gets generated at the next conflict. An example of this is when the solver was supposed to backtrack over a literal that would end up being unit propagated at level 0, but stays assigned at its current decision level when chronological backtracking is used. Additionally, chronological backtracking breaks the invarant that all literals appear on the trail in monotonically increasing order by decision level, which makes its implementation slightly more complicated.
